\documentclass[12pt,letterpaper,reqno]{amsart}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage[normalem]{ulem}
\usepackage{titlesec,bbm, hyperref}
\usepackage{spverbatim} 
\usepackage{esvect}
\usepackage{geometry}
\usepackage{caption}
\usepackage{subcaption}
\geometry{letterpaper, portrait, margin=0.5in}

\newcommand{\R}{\mathbb R}

\begin{document}



\begin{center}\large{
    MATH 631\quad
    HW 1\quad
    Sumanth Ravipati\quad
    January 28, 2019}
\end{center}
\vspace{.25in}

\begin{enumerate}
\item[1.] Let X be a set. Define a collection of subsets of X to be \textit{closed}. Assume that
\begin{enumerate}
    \item If $A$ and $B$ are closed sets, then $A \cup B$ is closed.
    \item If $\{A_i | i \in I \}$ is a collection of closed sets, then $\cap_{i \in I} A_i$ is closed.
    \item $\emptyset$ and $X$ are closed. \newline
\end{enumerate}
Prove that if we define open sets on $X$ to be the complement of closed sets, then this makes $X$ into a topological space (as we defined on the first day).
\begin{flushleft}

\end{flushleft}
Conversely, if $X$ is a topological space, show that closed sets (with respect to the topology on $X$) satisfy the above three axioms.
\begin{flushleft}

\end{flushleft}
\item[2.] Let $X$ be any set. Define $d : X \times X \rightarrow \R$ by $d(x,y) = 1$ if $x \not= y$ and $d(x,y) = 0$ if $x=y$. Prove that $d$ is a metric. Then prove that the topology induced from this metric is the discrete topology.
\begin{flushleft}

\end{flushleft}
\item[3.] Let $d(x,y) = |x-y|$ for $x,y \in \R$. Prove that this defines a metric on $\R$. We call the induced topology from this metric the \textit{standard} topology on $\R$. Whenever we do not stipulate a topology on $\R$ we mean the standard one.
\begin{flushleft}

\end{flushleft}
\item[4.] Let $(X,d)$ be a metric space and $x_0 \in X$. Prove that $f : X \rightarrow \R$ given by $f(x) = d(x,x_0)$ is continuous.
\begin{flushleft}

\end{flushleft}
\item[5.] Let $(X,d)$ be a metric space, and let $A \subset X$. Define $d_A : X \rightarrow \R$ by $d_A(x) = \inf\{d(x,y) | y \in A\}$. The value $d_A(x)$ is called the \textit{distance from $x$ to} $A$. Prove that $d_A$ is continuous.
\begin{flushleft}

\end{flushleft}
\end{enumerate}
\end{document}